Spin canonical divisors of log stacky curves

### 2015-06-21 stacky_write_up/spin_cring.tex
add base cases, exception cases, concl, saturation
* also added tables
* also shortened label names
* added subscript to S(i)

\usepackage{amsmath}
\usepackage{amsxtra}
\usepackage{amscd}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{eucal}
\usepackage[all]{xy}
\usepackage{graphicx}
\usepackage{tikz-cd}
\usepackage{mathrsfs}
\usepackage{subfiles}
\usepackage{mathpazo}
\usepackage{euler}
\usepackage{hyperref}
\usepackage{color}

@@

\usepackage{hyperref}
  \hypersetup{colorlinks=true,citecolor=blue}

@@

\subsection{Main Results}
These are our results.
\begin{thm}
\label{thm:g_0_generators_relations}
Let $(\sx,\Delta, L)$ be a tame log spin curve over a perfect field $k$, so
that $\sx$ has signature $\sigma = (0;e_1,\ldots, e_r;\delta)$. Then the
canonical ring

\begin{align*}
	R(\sx, \Delta, L) = \bigoplus_{d=0}^\infty H^0(\sx,\sco(L)^{\otimes d})
\end{align*}

\noindent
is generated as a $k$-algebra by elements of degree at most $e = 
\max(5,e_1,\ldots, e_r)$ with relations in degree at most $2e$, so 
long as $\sigma$ does not lie in a finite list of exceptions. \todo{
Include reference to the table with this list of exceptions.}
\end{thm}
\begin{proof}

@@

In this section, we will prove that if $(\sx , 0, L)$ is a spin curve,
so that $\sx$ has signature $\sigma = (0; e_1, \ldots , e_r; 0)$, then
$R(\sx , 0, L)$ is generated in degree at most $e := \max(5, e_1, \ldots,
e_r)$ with relations in degree at most $2e$, so long as $\sigma$ does
not lie in a finite list of exceptions. This is proven in
Subsection ~\ref{ssec:g_0_conclusion}. The proof is a rather
involved induction. In Subsections ~\ref{ssec:g_0_ram_pts},
~\ref{ssec:g_0_ram_orders} we prove inductive theorems, and in
Subsection ~\ref{ssec:g_0_base} we check the base cases of the
induction. Finally, in Subsection ~\ref{ssec:g_0_exceptional} we
list out the exceptional cases.

\ssec{Saturation}
\label{ssec:g_0_saturation}

@@

\ssec{Inductive theorem: Increasing the Ramification Orders}
\label{ssec:g_0_ram_orders}
In this section, we aim to prove the inductive theorem, Theorem
~\ref{thm:ramification_order_induction}. Under fairly general
conditions, this theorem will allow us to show that if Theorem ~\ref
{thm:ramification_order_induction} holds for a curve with signature
$(0;e_1',\ldots, e_r';0)$ then it also holds for a curve with
signature $(0;e_1,\ldots, e_r;0)$ with $e_i' - e_i \in \{0,2\}$.

@@

\begin{enumerate}
	\item[(Ad-i)] Suppose for all $i \in J$, there exists $y_{i,e_i'} 
		\in R'$ so that $R'$ admits a presentation
		\begin{align*}
			R' \cong \left( k[x] \otimes k[y_{i,e_i'}]_{i \in J} \right)/I'
		\end{align*}
		so that for all $i \in J$ we have $\deg y_{e_i'} = e_i'$ and $-ord_{
		Q_i}(y_{i,e_i'}) = \frac{e_i'-1}{2}$.
	\item[(Ad-ii)] For all $i \in J$, any generator $z \neq y_{i,e_i'}$
		of $R'$ satisfies
		\begin{align*}
			\frac{- ord_{Q_i}(z)}{\deg z} < \subhalf {e_i'}.
		\end{align*}
	\item[(Ad-iii)] For all $i \in J$ we have
		\begin{align*}
			\deg \lfloor e_i L \rfloor \geq \# S(i)
		\end{align*}
		where
		\begin{align*}
			S(i) := \{j \in J : j \neq i \text{ and } e_j'+2d \mid e_i -
			e_j' \text{ for some } d \geq 0\}
		\end{align*}
\end{enumerate}
\end{defn}

@@

\begin{thm}
\label{thm:ramification_order_induction}
Suppose $(\sx', 0, L')$ is a tame separably rooted log stacky spin
curve and $(\sx', J)$ is admissible with generators $y_{i,e_i'} \in
R' = R_{L'}$ as in (Ad-i) of Definition ~\ref{defn:admissible}.
Then the following are true:
\begin{enumerate}
	\item[(a)] For all $i \in J$, there exists $y_{i,e_i} \in
		H^0(\sx, e_i(K_\sx))$ so that
		\begin{align*}
			-ord_{Q_i}(y_{i,e_i}) = \frac{e_i-1}{2}
		\end{align*}
		and
		\begin{align*}
			\frac{-\ord_{Q_j}(y_{i,e_i'})}{y_{i,e_i}} \leq \subhalf{e_j'}-
			\frac{1}{\deg(y_{i,e_i'})e_j'}
		\end{align*}
		for all $j \in J$ with $j \neq i$.
	\item[(b)] The elements $y_{i,e_i'}^ay_{i,e_i}^b,$ with $a \geq 0,
		b > 0$ span $R$ over $R'$ and the elements $y_{i,e_i}$ minimally
		generate $R$ over $R'$.
	\item[(c)] Equip $k[y] = k[y_{i,e_i}]_{i \in J}$ and $k[x]$ with
		any graded monomial order and $k[y,x] = k[y] \otimes k[x]$ with
		block order. Let $R = k[y,x]/I.$ Then,
		\begin{align*}
			in_\prec(I) = in_\prec(I')k[x,y] + \langle y_{i,e_i}z \mid z 
			\neq y_{i,e_i},y_{i,e_i'} \rangle
		\end{align*}
		as $z$ ranges over all generators of $R$.
	\item[(d)] Suppose $in_\prec(I')$ is minimally generated by
		products of two monomials and that for all $i \in J$, we have $e
		_i > \deg z$ for any generator $z$ of $R'.$ Then any set of
		minimal generators for $I'$ together with any set of relations as in 
		part (c) minimally generate $I$.
	\item[(e)] For any $W \subset J,$ the set $(\sx,W)$ is admissible.
\end{enumerate}
\end{thm}

@@

\ssec{Base Cases}
\label{ssec:g_0_base}
Cases to induct on are here.

\ssec{Exceptional Cases}
\label{ssec:g_0_exceptional}
These are exceptional cases.

@@

>>>>

\usepackage{amsmath}
\usepackage{amsxtra}
\usepackage{amscd}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{eucal}
\usepackage[all]{xy}
\usepackage{graphicx}
\usepackage{tikz-cd}
\usepackage{mathrsfs}
\usepackage{subfiles}
\usepackage{mathpazo}
\usepackage{euler}
\usepackage{hyperref}
\usepackage{color}
\usepackage{longtable}

@@

%\usepackage{hyperref}
  %\hypersetup{colorlinks=true,citecolor=blue}

@@

\subsection{Main Results}
These are our results.
\begin{thm}
\label{thm:g_0_generators_relations}
Let $(\sx, \Delta, L)$ be a tame log spin curve over a perfect field $k$, so
that $\sx$ has signature $\sigma = (0; e_1, \ldots, e_r; \delta)$. Then the
canonical ring

\begin{align*}
	R(\sx, \Delta, L) = \bigoplus_{d = 0}^\infty H^0(\sx, \sco(L)^{\otimes d})
\end{align*}

\noindent
is generated as a $k$-algebra by elements of degree at most $e = 
\max(5,e_1,\ldots, e_r)$ with relations in degree at most $2e$, so 
long as $\sigma$ does not lie in a finite list of exceptions. \todo{
Include reference to the table with this list of exceptions.}
\end{thm}

@@

In this section, we will prove that if $(\sx , 0, L)$ is a spin curve,
so that $\sx$ has signature $\sigma = (0; e_1, \ldots , e_r; 0)$, then
$R(\sx , 0, L)$ is generated in degree at most $e := \max(5, e_1, \ldots,
e_r)$ with relations generated in degree at most $2e$, so long as $\sigma$
does not lie in a finite list of exceptions. This is proven in
Subsection ~\ref{ssec:g_0_conclusion}. The proof is a rather
involved induction. In Subsections ~\ref{ssec:g_0_ram_pts},
~\ref{ssec:g_0_ram_orders} we prove inductive theorems, and in
Subsection ~\ref{ssec:g_0_base} we check the base cases of the
induction. Finally, in Subsection ~\ref{ssec:g_0_exceptional} we
list out the exceptional cases.

\ssec{Saturation}
\label{ssec:g_0_saturation}
Here we compute saturation numbers. That is, given a list of
ramification orders, with $g = 0,\delta = 0$ understood, we have
the following saturations of the canonical divisor, as defined in 7.
2 of VZB \cite{vzb:stacky}. Exceptional cases are listed first
and generic cases follow.

\todo{prove}

\begin{longtable}
	{| c | c |}
	\hline
	Signature & Saturation \\
	\hline
	\hline

	$(0; 3, 3, 3; 0)$ & $\infty$ \\	\hline

	$(0; 3, 3, 5; 0)$ & $18$ \\	\hline

	$(0; 3, 3, 7; 0)$ & $12$ \\	\hline

	$(0; 3, 3, 9; 0)$ & $12$ \\	\hline

	$(0; 3, 5, 5; 0)$ & $8$ \\	\hline

	$(0; 5, 5, 5; 0)$ & $8$ \\	\hline

	$(0; 3, 3, 3, 3; 0)$ & $6$ \\	\hline

	\hline

	$(0; 3, 3, k; 0)$ & $9$ \\	\hline

	$(0; a, b, c; 0)$ & $5$ \\	\hline

	$(0; e_1, \ldots, e_r; 0)$ & $3$ \\	\hline
\end{longtable}

\begin{center}
\textbf{Table (I)}: Genus 0 saturation
\end{center}

@@

\ssec{Inductive Theorem: Increasing the Ramification Orders}
\label{ssec:g_0_ram_orders}
In this section, we aim to prove the inductive theorem, Theorem
~\ref{thm:ram_order_ind}. Under fairly general
conditions, this theorem will allow us to show that if Theorem ~\ref
{thm:ram_order_ind} holds for a curve with signature
$(0;e_1',\ldots, e_r';0)$ then it also holds for a curve with
signature $(0;e_1,\ldots, e_r;0)$ with $e_i' - e_i \in \{0,2\}$.

@@

\begin{enumerate}
	\item[(Ad-i)] Suppose for all $i \in J$, there exists $y_{i,e_i'} 
		\in R'$ so that $R'$ admits a presentation
		\begin{align*}
			R' \cong \left( k[x] \otimes k[y_{i,e_i'}]_{i \in J} \right)/I'
		\end{align*}
		so that for all $i \in J$ we have $\deg y_{e_i'} = e_i'$ and $-ord_{
		Q_i}(y_{i,e_i'}) = \frac{e_i'-1}{2}$.
	\item[(Ad-ii)] For all $i \in J$, any generator $z \neq y_{i,e_i'}$
		of $R'$ satisfies
		\begin{align*}
			\frac{- ord_{Q_i}(z)}{\deg z} < \subhalf {e_i'}.
		\end{align*}
	\item[(Ad-iii)] For all $i \in J$ we have
		\begin{align*}
			\deg \lfloor e_i L \rfloor \geq \# S_{\sigma, J}(i)
		\end{align*}
		where
		\begin{align*}
			S_{\sigma, J}(i) := \{j \in J : j \neq i \text{ and } e_j'+2d
			\mid e_i - e_j' \text{ for some } d \geq 0\}
		\end{align*}
\end{enumerate}
\end{defn}

@@

\begin{thm}
\label{thm:ram_order_ind}
Suppose $(\sx', 0, L')$ is a tame separably rooted log stacky spin
curve and $(\sx', J)$ is admissible with generators $y_{i,e_i'} \in
R' = R_{L'}$ as in (Ad-i) of Definition ~\ref{defn:admissible}.
Then the following are true:
\begin{enumerate}
	\item[(a)] For all $i \in J$, there exists $y_{i,e_i} \in
		H^0(\sx, e_i(K_\sx))$ so that
		\begin{align*}
			-ord_{Q_i}(y_{i,e_i}) = \frac{e_i-1}{2}
		\end{align*}
		and
		\begin{align*}
			\frac{-\ord_{Q_j}(y_{i,e_i'})}{y_{i,e_i}} \leq \subhalf{e_j'}-
			\frac{1}{\deg(y_{i,e_i'})e_j'}
		\end{align*}
		for all $j \in J$ with $j \neq i$.
	\item[(b)] The elements $y_{i,e_i'}^ay_{i,e_i}^b,$ with $a \geq 0,
		b > 0$ span $R$ over $R'$ and the elements $y_{i,e_i}$ minimally
		generate $R$ over $R'$.
	\item[(c)] Equip $k[y] = k[y_{i,e_i}]_{i \in J}$ and $k[x]$ with
		any graded monomial order and $k[y,x] = k[y] \otimes k[x]$ with
		block order. Let $R = k[y,x]/I.$ Then,
		\begin{align*}
			in_\prec(I) = in_\prec(I')k[x,y] + \langle y_{i,e_i}z \mid z 
			\neq y_{i,e_i},y_{i,e_i'} \rangle
		\end{align*}
		as $z$ ranges over all generators of $R$.
	\item[(d)] Suppose $in_\prec(I')$ is minimally generated by
		products of two monomials and that for all $i \in J$, we have $e
		_i > \deg z$ for any generator $z$ of $R'.$ Then any set of
		minimal generators for $I'$ together with any set of relations as in 
		part (c) minimally generate $I$.
	\item[(e)] For any $W \subset J,$ the set $(\sx,W)$ is admissible.
\end{enumerate}
\end{thm}

@@

\ssec{Base Cases}
\label{ssec:g_0_base}
Now we have an inductive statement on the number of ramified
points in Proposition ~\ref{prop:sat_three_induction} and
an inductive statement on ramification orders in Theorem
~\ref{thm:ram_order_ind}. After verifying that explicit pairs
$(\sx, J)$ of stacky curves $\sx$ with particular signatures and
particular choices of $J$ are admissible, the application of the
two inductive statements demonstrate that the spin canonical ring
$R(\sx , 0, L)$ is generated in degree at most $e := \max(5, e_1,
\ldots, e_r)$ with relations generated in degree at most $2e$ for
$\sx$ with any signature $\sigma = (0; e_1, \ldots , e_r; 0)$ outside
of a finite set of exceptional cases.

\begin{lem}
\label{lem:g_0_admissible_cases}
Let $(\sx', \Delta, L')$ be a tame log spin curve over a perfect field $k$,
so that $\sx$ has signature $\sigma = (0; e_1, \ldots, e_r; 0)$.
Then $R(\sx', \Delta, L')$ is generated as a $k$-algebra by elements of
degree at most $e = \max(5,e_1,\ldots, e_r)$ with relations in degree at
most $2e$ and $(\sx', J)$ is admissible for the following cases:

\begin{enumerate}
	\item[(a)] $\sigma = (0; 3, 3, 11; 0), J = \{3\}$
	\item[(b)] $\sigma = (0; 3, 5, 9; 0), J = \{2, 3\}$
	\item[(c)] $\sigma = (0; 5, 5, 7; 0), J = \{1, 2, 3\}$
	\item[(d)] $\sigma = (0; 3, 3, 3, 5; 0), J = \{1, 2, 3, 4\}$
	\item[(e)] $\sigma = (0; 3, 3, 3, 3, 3; 0), J = \{1, 2, 3, 4, 5\}$
\end{enumerate}
\end{lem}

\begin{proof}
For each case, computations using a combination of Riemann-Roch and
comparing pole orders at each ramification point $P_i$ gives the
generator and relation degrees. Note that (Ad-i) always holds.
In each case we can also pick a set of generators such that
(Ad-ii) holds. \todo{justification for these}

The following table demonstrates that $R(\sx', \Delta, L')$ is
generated as a $k$-algebra by elements of degree at most $e$
with relations in degree at most $2e$ for each case.

\begin{longtable}
	{| c || c | c | c |}
	\hline
	Case & Generator Degrees & Minimal Relation Degrees & $e$\\
	\hline
	\hline

	(a) & \{3, 7, 9, 11\} & \{14, 18\} & 11\\	\hline

	(b) & \{3, 5, 7, 9\} & \{12, 14\}	& 9\\ \hline

	(c) & \{3, 5, 5, 7\} & \{10, 12\}	& 7\\ \hline

	(d) & \{3, 3, 4, 5\} & \{8,	9\} & 5\\ \hline

	(e) & \{3, 3, 3, 4,	4, 5\} & \{6, 7, 7, 8, 8, 8, 9, 9, 10\} & 5\\ \hline
\end{longtable}

\begin{center}
\textbf{Table (II)}: Genus 0 noneffective base cases
\end{center}

Furthermore, each case satisfies (Ad-iii) as demonstrated below.

\begin{enumerate}
	\item[(a)] $(\deg \lfloor e_3 \rfloor, \# S_{(\sigma, J)}(3)) = (1, 0)$

	\item[(b)] $(\deg \lfloor e_2 \rfloor, \#S_{(\sigma, J)}(2)) = (1, 0)$ \\
		$(\deg \lfloor e_3 \rfloor, \#S_{(\sigma, J)}(3)) = (1, 0)$

	\item[(c)] $(\deg \lfloor e_1 \rfloor, \#S_{(\sigma, J)}(1)) = (2, 1)$ \\
		$(\deg \lfloor e_2 \rfloor, \#S_{(\sigma, J)}(2)) = (2, 1)$ \\
		$(\deg \lfloor e_3 \rfloor, \#S_{(\sigma, J)}(3)) = (1, 0)$

	\item[(d)] $(\deg \lfloor e_i \rfloor, \#S_{(\sigma, J)}(i)) = (3, 1)$
		for $i \neq 4$ \\
		$(\deg \lfloor e_4 \rfloor, \#S_{(\sigma, J)}(4))	= (2, 0)$.

	\item[(e)] $(\deg \lfloor e_i \rfloor, \#S_{(\sigma, J)}(1)) = (5, 0)$
		for all $i \in J$.
\end{enumerate}

Finally, each case also satisfies the property that $in_\prec(I')$ is
minimally generated by products of two monomials and that for all
$i \in J$, we have $e_i > \deg z$ for any generator $z$ of $R'.$
\todo{prove}
\end{proof}

Now we can induct on cases (a)-(e) of \ref{lem:g_0_admissible_cases}
to obtain bounds on the minimal generator and relation degrees for
canonical rings of curves in the genus 0 case.

Using the results of Proposition ~\ref{prop:sat_three_induction}
and Theorem ~\ref{thm:ram_order_ind}, we can inductively raise
the ramification orders and number of ramified points in all of the
cases in Lemma ~\ref{lem:g_0_admissible_cases} to obtain the following
result.

\begin{cor}
\label{cor:g_0_gen_rel_base_cases}
Let $(\sx, \Delta, L)$ be a tame log spin curve over a perfect field $k$, so
that $\sx$ has signature $\sigma = (0; e_1, \ldots, e_r; 0)$. Then the
canonical ring

\[
	R(\sx, 0, L) = \bigoplus_{d = 0}^\infty H^0(\sx, \sco(L)^{\otimes d})
\]

\noindent
is generated as a $k$-algebra by elements of degree at most $e := 
\max(5, e_1, \ldots, e_r)$ with relations in degree at most $2e$, so 
long as $\sigma \notin S$ where

\begin{align*}
	S &:= \{(0; 3, 3, k; 0) : 3 \leq k \leq 9 \text{ odd}\} \\
		&\cup \{(0; 3, b, c; 0) : 5 \leq b \leq c \leq 7\} \\
		&\cup \{(0; 5, 5, 5; 0), (0; 3, 3, 3, 3; 0)\}
\end{align*}

\noindent
is a finite set.
\end{cor}

\begin{proof}
A process of induction raising the ramification orders of
the ramified points of the spin curves corresponding to Lemma
~\ref{lem:g_0_admissible_cases} cases (a)-(c) via Theorem
~\ref{thm:ram_order_ind} will demonstrate the desired generator and
relation degree results for curves with three ramified points.
Similarly, this induction process on Lemma
~\ref{lem:g_0_admissible_cases} case (d) will give the desired
result for curves with four ramified points excluding signature
$(0; 3, 3, 3, 3; 0)$. A process of induction increasing the
number of ramified points via Proposition
~\ref{prop:sat_three_induction} on Lemma ~\ref{lem:g_0_admissible_cases}
case (e) combined with the induction process on ramification orders
will yield the result for curves with at least five ramified points.
\todo{Detailed description of how the inductive lemmata yield e and 2e}
\end{proof}

\ssec{Exceptional Cases}
\label{ssec:g_0_exceptional}
Here we present the explicit generators and relations of the remaining
cases given by signatures in the finite set

\begin{align*}
	S &:= \{(0; 3, 3, k; 0) : 3 \leq k \leq 9 \text{ odd}\} \\
		&\cup \{(0; 3, b, c; 0) : 5 \leq b \leq c \leq 7\} \\
		&\cup \{(0; 5, 5, 5; 0), (0; 3, 3, 3, 3; 0)\}
\end{align*}

\noindent
and in particular describe the only exceptions to the $e$ and $2e$
bounds on the generator degree and relation degree, respectively,
in the genus 0 case.

\begin{longtable}
	{| c || c | c | c |}
	\hline
	Signature & Generator Degrees & Minimal Relation Degrees & $e$ \\
	\hline
	\hline

	$(0; 3, 3, 3; 0)$ & $\{3\}$ & $\emptyset$ & $3$ \\	\hline

	$(0; 3, 3, 5; 0)$ & $\{3, 10, 15\}$ & $\{30\}$ & $5$ \\	\hline

	$(0; 3, 3, 7; 0)$ & $\{3, 7, 12\}$ & $\{24\}$ & $7$ \\	\hline

	$(0; 3, 3, 9; 0)$ & $\{3, 7, 9\}$ & $\{21\}$ & $9$ \\	\hline

	$(0; 3, 5, 5; 0)$ & $\{3, 5, 10\}$ & $\{20\}$ & $5$ \\	\hline

	$(0; 3, 5, 7; 0)$ & $\{3, 5, 7\}$ & $\{17\}$ & $7$ \\	\hline

	$(0; 5, 5, 5; 0)$ & $\{3, 5, 5\}$ & $\{15\}$ & $5$ \\	\hline

	$(0; 3, 3, 3, 3; 0)$ & $\{3, 3, 4\}$ & $\{12\}$ & $3$ \\	\hline

	\hline

	$(0; 3, 7, 7; 0)$ & $\{3, 5, 7, 7\}$ & $\{10, 14\}$ & $7$ \\	\hline
\end{longtable}

\begin{center}
\textbf{Table (III)}: Genus 0 exceptional cases
\end{center}

\begin{rem}
These cases, apart from $(0; 3, 7, 7; 0)$, give all of the
exceptions to the $e$ and $2e$ bounds on the generator and relation
degree. Notice that each of these exceptional cases, apart from
$(0; 3, 5, 7; 0)$, corresponds to a signature with non-generic
saturation (i.e. saturation not equal to 3, 5, or 9 as seen
in the table in Subsection ~\ref{ssec:g_0_saturation}). Exceptional
saturation can be viewed as ``forcing'' generators (and thus
relations) in higher degrees than in the generic case.
\end{rem}

@@

\begin{thm}
\label{thm:g_0_generators_relations_rep}
Let $(\sx, \Delta, L)$ be a tame log spin curve over a perfect field $k$, so
that $\sx$ has signature $\sigma = (0; e_1, \ldots, e_r; \delta)$. Then the
canonical ring

\begin{align*}
	R(\sx, \Delta, L) = \bigoplus_{d = 0}^\infty H^0(\sx, \sco(L)^{\otimes d})
\end{align*}

\noindent
is generated as a $k$-algebra by elements of degree at most $e = 
\max(5,e_1,\ldots, e_r)$ with relations in degree at most $2e$, so 
long as $\sigma$ does not lie in a finite list of exceptions. \todo{
Include reference to the table with this list of exceptions.}
\end{thm}
